
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

load all packages used in this exploratory analysis
```{r loding pacakge and set the global functionality}
library(knitr)
library(dplyr)
library(ggplot2)
opts_chunk$set(echo = TRUE)
setwd("F:/Data Science/Course Eara/Course-5 Reproducable Research/Week-2/Project")
```
## Loading and preprocessing the data
```{r Loading data}
# load data
data_row <- read.csv('activity.csv')

# remove NA in data
data <- data_row[ with (data_row, { !(is.na(steps)) } ), ]

# print out first 20 rows
head(data,20)
```

## What is mean total number of steps taken per day?

```{r mean total number of Steps Taken}
by_day <- group_by(data, date)
steps_by_day <- summarise(by_day, total = sum(steps))
steps_by_day
```
```{r creating a histogram}
hist(steps_by_day$total, main="Histogram of total number of steps per day", 
     xlab="Total number of steps in a day")
```

## What is the average daily activity pattern?


```{r activity Patren}
# preprocessing data for plot
steps_by_interval <- aggregate(steps ~ interval, data, mean)

# create a time series plot 
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```

Finding maximum Steps in a Day
```{r finding max steps in a day}
# find row with max of steps
max_steps_row <- which.max(steps_by_interval$steps)

# find interval with this max
steps_by_interval[max_steps_row, ]
```


###Imputing missing values
```{r inputing missing Values}
sum(is.na(data_row))
```

##Total number of rows with NA's is `r sum(is.na(data_row))'
packed Replacing NA's with the mean for that 5-minute interval

```{r replacing NA with the mean for that 5-minute interval}
data_imputed <- data_row
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
    interval_value <- data_imputed$interval[i]
    steps_value <- steps_by_interval[
      steps_by_interval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
```
I've created new data set data_no_na which equals to data_row but without NA's. All NA's are replaced with mean of 5-minute interval.
```{r }
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
head(df_imputed_steps_by_day)
```

```{r }

hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
```

```{r}
# get mean and median of imputed data
mean(df_imputed_steps_by_day$steps)
median(df_imputed_steps_by_day$steps)
mean(steps_by_day$total)
median(steps_by_day$total)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data_imputed['type_of_day'] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
# convert type_of_day from character to factor
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)

# calculate average steps by interval across all days
df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, data_imputed, mean)

# creat a plot
qplot(interval, 
      steps, 
      data = df_imputed_steps_by_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
  facet_wrap(~ type_of_day, ncol = 1)
```